---
title: "Task 12"
author: "matthewchapman"
date: "2022-10-06"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning=FALSE)
```

### Libraries
```{r}
library(dplyr)
library(ggplot2)
library(tidyr)
library(lubridate)
library(downloader)
library(readr)
library(haven)
library(readxl)
library(forcats)
library(knitr)
library(tidyverse)
```

## Read in 5 files as temp files
```{r}
rds <- tempfile()
download.file("https://github.com/WJC-Data-Science/DTS350/raw/master/Dart_Expert_Dow_6month_anova/Dart_Expert_Dow_6month_anova.RDS", rds, mode = "wb")
rds <- readRDS(rds)

csv <- tempfile()
download.file("https://raw.githubusercontent.com/WJC-Data-Science/DTS350/master/Dart_Expert_Dow_6month_anova/Dart_Expert_Dow_6month_anova.csv", csv, mode = "wb")
csv <- read_csv(csv)

dta <- tempfile()
download.file("https://github.com/WJC-Data-Science/DTS350/raw/master/Dart_Expert_Dow_6month_anova/Dart_Expert_Dow_6month_anova.dta", dta, mode = "wb")
dta <- read_dta(dta)

sav <- tempfile()
download.file("https://github.com/WJC-Data-Science/DTS350/raw/master/Dart_Expert_Dow_6month_anova/Dart_Expert_Dow_6month_anova.sav", sav, mode = "wb")
sav <- read_sav(sav)

xl <- tempfile()
download.file("https://github.com/WJC-Data-Science/DTS350/raw/master/Dart_Expert_Dow_6month_anova/Dart_Expert_Dow_6month_anova.xlsx", xl, mode = "wb")
xl <- read_xlsx(xl)
```

### Confirm all 5 files are all_equal()
```{r}
all_equal(rds,csv,dta,convert = TRUE)
all_equal(rds,sav,xl,convert = TRUE)
```

## Performance Visualizations
### Box Plot
```{r}
ggplot(xl,mapping = aes(x = variable,y = value, color = variable,fill=variable)) +
  geom_boxplot(alpha=3/5) +
  labs(x = "Variable",y="Value",title="Box Plots with Each Variable and their Values")
```
### Jittered Returns
```{r}
ggplot(xl, mapping = aes(x=variable,y=value)) +
  geom_jitter(aes(color=variable)) +
  labs(x = "Variable",y="Value",title="Jitter Plot with Each Variable and their Values")
```
### Average Return
```{r}
ggplot(xl,mapping = aes(x=contest_period,y=value,color = variable,group = variable)) +
  geom_point() +
  geom_smooth(method = "lm",color = "black") +
  facet_wrap(~ variable)
```

## Tidy the Data
```{r}
months <- c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December")

data <- xl %>%
  mutate(contest_period = str_replace_all(contest_period, "19", "_19"),
         contest_period = str_replace_all(contest_period, 'Dec._1993', 'December_1993'),
         contest_period = str_replace_all(contest_period, 'Febuary', 'February')) %>%
  separate(contest_period, into = c("Month", 'year_end'),sep="-") %>%
  separate(year_end, into = c("Month_end", 'Year_end'),sep="_") %>%
  select(-Month) %>%
  mutate(Month_end = factor(Month_end, levels = months))
data
```
### Save tidied data as rds object
```{r}
saveRDS(data, file = "data.rsd")
```

## Visualize 6-month returns
```{r}
ggplot(data = data, aes(x = Month_end, y = value,group = Year_end)) +
  geom_jitter(aes(color = Month_end)) +
  geom_smooth(method = "lm",color="black") +
  facet_wrap(~Year_end) +
  labs(x="Last Month",y = "Value",title = "Value at the ending of Each Month by Year",color = "Last Month")
```

## Match table below
```{r}
data1 <- data %>%
  filter(variable == "DJIA") %>%
  pivot_wider(names_from = "Year_end", values_from = "value") %>%
  arrange(factor(Month_end, levels = months))
data1
```

This visualization shows the value for each month, for each variable, faceted by year. There isn't too much that can be taken from this visualization, as there are not many noticeable trends. It appears that the months of May - Septembers are the furthest away from the line in a lot of cases.