---
title: "Task4"
author: "matthewchapman"
date: "2022-09-08"
output: 
  html_document:
    keep_md: TRUE
---
#Convert iris data to a tibble called iris_data
```{r}
library("tibble")
library("dplyr")
data("iris")
class(iris)

iris_data <- as_data_frame(iris) 
class(iris_data)
iris_data
```


#Arrange the iris data by Sepal.Length and display the first ten rows.
```{r}
arrange(iris_data, Sepal.Length)
```


#Select the Species and Petal.Width columns and put them into a new data set called testdat.
```{r}
testdat <- select(iris, Species, Petal.Width)
testdat
```


#Create a new table called species_mean that has the mean for each variable for each Species.
```{r}
species_mean <- group_by(iris, Species)
species_mean <- summarise(species_mean, Petal.Length = mean(Petal.Length, na.rm = TRUE), Petal.Width = mean(Petal.Width, na.rm = TRUE), Sepal.Length = mean(Sepal.Length, na.rm = TRUE), Sepal.Width = mean(Sepal.Width, na.rm = TRUE))
species_mean
```


#Create a data frame called iris_min with only the observations with at least 3 for Sepal.Width and not 2.5 for Petal.Width.
```{r}
iris_min <- filter(iris_data, Sepal.Width >= 3, Petal.Width != 2.5)
iris_min
```


#Create a data frame called iris_size with a column Sepal.Size
```{r}
iris_size <- 
  mutate(iris_data, Sepal.Size= case_when(
    Sepal.Length < 5 ~ "small",
    Sepal.Length >= 5 & Sepal.Length <6.5 ~ "medium",
    Sepal.Length >= 6.5 ~ "large"
  )
  )
iris_size
```


#Create a data frame called iris_rank which ranks all of the data by Petal.Length, where a ranking of 1 is the largest Petal.Length and order by ranking.
```{r}
iris_rank <-
  mutate(iris_data, ranking = max_rank(Petal.Length))
iris_rank
```


#Read about the ?summarize_all() function and get a new table with the means and standard deviations for each species.
```{r}
iris3 <- iris %>%
  group_by(Species) %>%
  summarise_all(list(mean, sd))
iris3
```


#Questions:
#1. What part of the country has the cheapest cost of living and has the most major cities within 200 miles?
#2. Where is the best place to live working in tech?
#3. Is there going to be another bad COVID outbreak?
#4. Are major corporations loss prevention systems worth it?