---
title: "Task 14"
author: "matthewchapman"
date: "2022-10-16"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE)
```

### Libraries
```{r}
library(tidyverse)
library(downloader)
library(dplyr)
library(tidyr)
library(stringi)
library(stringr)
library(ggplot2)
```

### Import Bible Data
```{r}
bible <- read_csv("https://raw.githubusercontent.com/WJC-Data-Science/DTS350/master/bible.csv")
head(bible)
```


## Average verse length (Old Testament vs New Testament)
### Function that returns the average verse length of a data set
```{r}
length.function <- function(df, f) {
  word_length <- vector("integer",length(df[[f]]))
  for (i in seq_along(word_length)) {
    word_length[[i]] <- str_count(df[[f]][i],'\\w+')
  }
  word_avg <- sum(word_length)/length(df[[f]])
  word_avg
}
```
### Create Data sets to input into function
```{r}
Old_Test <- bible %>%
  filter(volume_title == "Old Testament")
Old_Test

New_Test <- bible %>%
  filter(volume_title == "New Testament")
New_Test
```
### Find Average Verse Length for each testament
```{r}
length.function(Old_Test,15)
length.function(New_Test,15)
```


## LORD Frequency (Old Testament vs New Testament)
```{r}
bible2 <- bible %>%
  select(volume_title,scripture_text) %>%
  mutate(lord_count = str_count(scripture_text, coll("lord",ignore_case = TRUE))) %>%
  group_by(volume_title) %>%
  summarise(LORD_count = sum(lord_count))
bible2

bible21 <- bible %>%
  select(volume_title,scripture_text) %>%
  mutate(lord_count = str_count(scripture_text, coll("lord",ignore_case = TRUE))) %>%
  group_by(volume_title) %>%
  summarise(LORD_count = mean(lord_count))
bible21
```


## Word count (by verse) for each book in New Testament
### Prepare Data for Visualization 
```{r}
bible3 <- bible %>%
  filter(volume_title == "New Testament") %>%
  select(book_title,verse_id,scripture_text) %>%
  mutate(word_count = str_count(scripture_text,'\\w+')) %>%
  group_by(book_title) %>%
  mutate(avg_word_count = mean(word_count)) 
bible3
```
### Plot
```{r}
ggplot(data=bible3, mapping = aes(x = book_title,y = word_count)) +
  geom_boxplot(aes(color = book_title)) +
  theme(axis.text.x=element_blank()) +
  labs(x = "", y = "Length of Verse (Word Count)", color = "Book Title", title = "Length of Bible Verses by Book (NT)")

```

For the first question, I found that the Old Testament has an average verse length of 26.4 words, which is longer compared to the New Testament at 22.7 words per verse. For the second question, I found that LORD is much more frequent in the Old Testament (7273 total), compared to the New Testament (736 total). The Old Testament also has an average of 0.314 "LORD"s per verse, and the New Testament has an average of 0.0925 "LORD"s per verse. For the last question, I created a box plot to show the distribution of length of verses for each book in the New Testament. The plot shown confirms that the average verse length is roughly 22-23. There is not much variance among each book, they are all in the same general area. Revelations is the only book that really sticks out to me, with the largest average length per verse.
