---
title: "CS9"
author: "matthewchapman"
date: "2022-11-03"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,warning = FALSE,message = FALSE)
```

#### Libraries
```{r}
library(dplyr)
library(timetk)
library(ggplot2)
library(tidyquant)
library(tidyverse)
library(dygraphs)
library(lubridate)
```

# SECTION 1
### Tickers Today
```{r}
tickers_today <- tq_get(c("CXW", "F", "GM", "JCP", "KR", "WDC", "NKE","T", "WDAY", "WFC", "WMT"),get = "stock.prices", from = today()-years(5))
head(tickers_today)
```

### Build Interactive Visualizations
#### Win/Loss Each Day
```{r}
tickers <- tickers_today %>%
  group_by(symbol) %>%
  mutate(starting = open[1]) %>%
  select(symbol,date,open,close,volume,starting) %>%
  mutate(value = starting - close) %>%
  select(symbol,date,value) %>%
  pivot_wider(names_from = symbol, values_from = value) %>%
  tk_xts(date_var = date)
head(tickers)
```
#### Value Each Day
```{r}
My_stockPrices <- tickers_today %>%
  select(symbol, date, adjusted) %>%
  pivot_wider(names_from = symbol, values_from = adjusted) %>%
  tk_xts(date_var = date)
head(My_stockPrices)
```
#### Graphs
```{r}
dygraph(tickers,main = "Stock Performance (Win/Loss ($))")


dygraph(My_stockPrices, main = "Stock Performance (Percentage)") %>%
  dyRebase(percent = TRUE) 
```


### Volume
#### Dataset with Volume
```{r}
My_volume <- tickers_today %>%
  select(symbol,date,volume)
head(My_volume)
```
#### Graphs
```{r}
ggplot(My_volume,mapping = aes(x = symbol,y=volume,fill=symbol)) +
  geom_boxplot()

ggplot(My_volume,mapping = aes(x = date,y=volume,fill=symbol)) +
  geom_col() +
  facet_wrap(~symbol)
```





# SECTION 2
### Enter number list of stocks
```{r}
#x <- c()
```
### Tickers Today
```{r}
#tickers_today1 <- tq_get(x,get = "stock.prices", from = today()-years(5))
#head(tickers_today1)
```

### Build Interactive Visualizations
#### Win/Loss Each Day
```{r}
#tickers1 <- tickers_today1 %>%
 # group_by(symbol) %>%
  #mutate(starting = open[1]) %>%
  #select(symbol,date,open,close,volume,starting) %>%
  #mutate(value = starting - close) %>%
  #select(symbol,date,value) %>%
  #pivot_wider(names_from = symbol, values_from = value) %>%
  #tk_xts(date_var = date)
#tickers1
```
#### Value Each Day
```{r}
#My_stockPrices1 <- tickers_today1 %>%
 # select(symbol, date, adjusted) %>%
  #pivot_wider(names_from = symbol, values_from = adjusted) %>%
  #tk_xts(date_var = date)
#My_stockPrices1
```
#### Graphs
```{r}
#dygraph(tickers1,main = "Stock Performance (Win/Loss ($)")


#dygraph(My_stockPrices1, main = "Stock Performance (Percentage)") %>%
 # dyRebase(percent = TRUE) 
```


### Volume
#### Dataset with Volume
```{r}
#My_volume1 <- tickers_today1 %>%
 # select(symbol,date,volume)
#My_volume1
```
#### Graphs
```{r}
#ggplot(My_volume1,mapping = aes(x = symbol,y=volume,fill=symbol)) +
 # geom_boxplot()

#ggplot(My_volume1,mapping = aes(x = date,y=volume,fill=symbol)) +
#  geom_col() +
#  facet_wrap(~symbol)
```

After looking at the performance in percentage interactive graph for one stock, I've determined that KR is the best choice, it has been performing well for the entire 5 year period and survived the pandemic. WDAY and WMT are also performing well at 2nd and 3rd place. CXW is performing the worst. I decided to make a win loss graphic as well because not all stock prices are equal. A 5% drop in a 1000 dollar stock is far worse than a 5% drop in a 100 dollar stock. For volume, I used a box plot and faceted bar chart to show the disribution of volume with each stock.