---
title: "Task11"
author: "matthewchapman"
date: "2022-10-03"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

#### Libararies
```{r}
library(downloader)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(RColorBrewer)
library(tidyr)
```

#### Download Data
```{r}
data <- read_csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/fandango/fandango_score_comparison.csv")
data
```

#### Tidy Data
```{r}
d1 <- data %>%
  select(FILM, RT_norm, RT_user_norm, Metacritic_norm, Metacritic_user_nom, IMDB_norm) %>%
  pivot_longer(c('RT_norm', 'RT_user_norm','Metacritic_norm','Metacritic_user_nom','IMDB_norm'),names_to = "Website",values_to = "Score") %>%
  group_by(Website) %>%
  summarise(Score = mean(Score)) %>%
  mutate(Website = fct_reorder(Website, Score)) %>%
  mutate(Type = case_when(
    Website == 'RT_norm' | Website == 'Metacritic_norm' ~ "Critic",
    Website == 'RT_user_norm' | Website == 'Metacritic_user_nom' | Website == 'IMDB_norm' ~ "User"
  ))
d1
```


#### Critic vs User Data
```{r}
critic <- d1 %>%
  filter(Website == "RT_norm" | Website == "Metacritic_norm") %>%
  summarise(Type,Score = mean(Score)) %>%
  group_by(Type,Score) %>%
  summarise()
critic

user <- d1 %>%
  filter(Website == "RT_user_norm" | Website == "Metacritic_user_nom" | Website == "IMDB_norm") %>%
  summarise(Type,Score = mean(Score)) %>%
  group_by(Type,Score) %>%
  summarise()
user
```

#### Plot 1
```{r}
plot1 <- ggplot(data = d1,mapping = aes(x = Type, y = Score, fill = Website)) +
  geom_col(position = "dodge") +
  geom_text(aes(x = 'Critic', y = 2.8, label = "Critic Average (2.99)",group = "Critic")) +
  geom_text(aes(x = 'User', y = 3.1, label = "User Average (3.27)")) +
  scale_fill_discrete(labels=c('Metacritic Critic', 'RT Critic', 'RT User','Metacritic User','IMDB User')) +
  geom_hline(aes(yintercept = Score), critic,linetype = "dashed") +
  geom_hline(aes(yintercept = Score), user)
plot1
```

#### First 20 rows of tidy data
```{r}
d2 <- data %>%
  select(FILM, RT_norm, RT_user_norm, Metacritic_norm, Metacritic_user_nom, IMDB_norm) %>%
  pivot_longer(c('RT_norm', 'RT_user_norm','Metacritic_norm','Metacritic_user_nom','IMDB_norm'),names_to = "Website",values_to = "Score")

d2 <- head(d2,n=20)
d2
```


The graph shows the average rating for the websites on a 5 star scale. Metacritic and Rotten Tomatoes both have a critic and user score, while IMDB only has user scores. I grouped them together by type and it turns out that users give a higher score than critics. The lowest average user score (RT User) is still higher than the highest critic score.