---
title: "Task20"
author: "matthewchapman"
date: "2022-10-28"
output: 
  html_document:
  keep_md: TRUE
---
  
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE)
```

### Libraries
```{r}
library(tidyverse) 
library(downloader)
library(dplyr)
library(tidyr)
library(stringi)
library(stringr)
library(ggplot2)
library(lubridate)
library(riem)
library(remotes)
library(devtools)
library(tidyverse)
library(ggplot2)
library(readr)
library(gapminder)
library(gganimate)
library(gifski)
```

I chose to go back to Case Study 8 for Task20. I decided to do a transition between each stores daily hours and sales. I think it's a great way to be interactive with the graph while still showing the important message.

### Import Sales Data
```{r}
sales <- read_csv("https://github.com/WJC-Data-Science/DTS350/raw/master/sales.csv")
head(sales)
tail(sales)
```

## Format Data for Analysis
```{r}
sales1 <- sales %>%
  with_tz(tz = "US/Mountain") %>%
  mutate(day = ceiling_date(Time, "day"),
         week = ceiling_date(Time, "week"),
         month = ceiling_date(Time, "month"),
         hour = ceiling_date(Time, "hour")) %>%
  filter(Name != "Missing")
head(sales1)
```

## Which Business is the best investment?
#### Hours of Operation
```{r}
ho2 <- sales1 %>%
  select(Name, Amount, hour) %>%
  mutate(hour = hour(hour)) %>%
  group_by(Name,hour) %>%
  summarise(Amount = sum(Amount)) %>%
  arrange(Name,hour)
head(ho2)

ggplot(ho2,mapping = aes(x = hour,y=Amount,fill=Name)) +
  geom_col() +
  transition_states(Name) +
  labs(x = "Hour of Day",y="Total Sales",title = "Total Sales Throughout Day for Each Store",subtitle = "Store: {closest_state}") +
  theme(legend.position = "none")

anim_save("FirstGif.gif")