---
title: "CS10"
author: "matthewchapman"
date: "2022-11-14"
output: 
  html_document:
    keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message = FALSE,warning = FALSE)
```

#### Libraries
```{r}
library(remotes)
library(devtools)
library(sf)
library(ggsflabel)
library(maps)
library(USAboundaries)
library(USAboundariesData)
library(tidyverse)
library(ggplot2)
library(readr)
```

#### Read in Permit and County Data
```{r}
permits <- read_csv("https://github.com/WJC-Data-Science/DTS350/raw/master/permits.csv")
head(permits)

county1 <- us_counties()
```

#### FIPS
```{r}
FIPS <- permits %>%
  mutate(state=as.character(state),stateFIPS=str_pad(state,2,pad="0")) %>%
  mutate(county=as.character(county),countyFIPS=str_pad(county,3,pad="0")) %>%
  mutate(geoid=paste0(stateFIPS,countyFIPS))
FIPS
```

#### Merge Data
```{r}
merge1 <- merge(FIPS,county1,by="geoid") %>%
  group_by(state_name,year) %>%
  summarise(across(value,sum)) %>%
  ungroup() %>%
  group_by(year) %>%
  mutate(country = mean(value))
merge1
```

#### US PLOT
```{r}
US <- ggplot(data = merge1,mapping=aes(x=year,y=value,color = state_name)) +
  geom_line() +
  geom_line(mapping = aes(x=year,y=country,size = 3)) +
  theme(legend.position = "none")
US
```

#### KS and MO Data
```{r}
KS <- merge1 %>%
  filter(state_name == 'Kansas')
KS
MO <- merge1 %>%
  filter(state_name == 'Missouri')
MO
```

#### KS PLOT
```{r}
KS1 <- ggplot(data = KS,mapping = aes(x = year,y = value)) +
  geom_line(color='red') +
  geom_point(color="blue") +
  labs(x = "Year", y = "Permite Amount",title = 'Kansas Permit Amounts')
KS1
```

#### MO PLOT
```{r}
MO1 <- ggplot(data = MO,mapping = aes(x = year,y = value)) +
  geom_line(color='gold') +
  geom_point(color="black") +
  labs(x = "Year", y = "Permite Amount",title = 'Missouri Permit Amounts')
MO1
```

#### KS vs MO PLOT
```{r}
KvM <- ggplot() +
  geom_line(data = MO,mapping = aes(x = year,y = value),color='gold') +
  geom_point(data = MO,mapping = aes(x = year,y = value),color="black") +
  geom_line(data = KS,mapping = aes(x = year,y = value),color='red') +
  geom_point(data = KS,mapping = aes(x = year,y = value),color="blue") +
  labs(x = "Year", y = "Permite Amount",title = 'Kansas vs Missouri Permit Amounts')
KvM
```

For all of my graphics I chose to use line graphs because they were the best choice for 2 numeric variables and the amount of data. To start things off I made a graph showing all the states over time with a average line. Most of the states showed a decline in the mid 1980's until the early 1990's. The numbers inlined steadily until just before the 2008 crash. I compared Kansas and Missouri because they are the two states I have lived in throughout my life and saw that they both have very similar line graphs in terms of trends. When I compared them by merging the graphs together, Missouri seemed to have larger numbers. However, I confirmed that the trends were the same.