---
title: "Task5"
author: "matthewchapman"
date: "2022-09-13"
output: 
  html_document:
    keep_md: TRUE
code_folding: 'hide'
---
Libraries
```{r}
library(tidyverse)
library(downloader)
library(readxl)
```

BillboardData
```{r}
download.file("https://raw.githubusercontent.com/DTS350Fall2022/ChapmanM/main/Week3/Task5/billboard-hits.csv","billboard-hits.csv")
BillboardData <- read_csv("billboard-hits.csv")
```

SoloData
```{r}
download.file("https://raw.githubusercontent.com/DTS350Fall2022/ChapmanM/main/Week3/Task5/solo-artist-followers.csv","solo-artist-followers.csv")
SoloData <- read_csv("solo-artist-followers.csv")
```

#Comment on Data Types:
Yes, Data imported correctly. I had to use the github raw URL to make it work. When I look at each data set, it is a tibble.


Remove singers who who do not have at least 6 times they had a top 100 hit.
```{r}
BillboardData1 <- 
  group_by(BillboardData, name) %>% 
  mutate(count = n()) %>%
  filter(count > 5) %>%
  select(name:peak_rank)
```


Find the corresponding data in the data set for the top singers’ bands.
```{r}
BillboardData2 <- 
  group_by(BillboardData1, band) %>% 
  mutate(count = n()) %>%
  filter(count > 5, !is.na(band))
```

Plot
```{r}

ggplot(data = BillboardData2) +
  geom_point(mapping = aes(x = peak_date, y = peak_rank, color = name)) +
  geom_line(aes(x= peak_date, y = peak_rank, color = name)) + 
  facet_wrap(~ band, nrow = 4, scales = "free")
```

TITANIC
```{r}
download.file("https://raw.githubusercontent.com/rashida048/Datasets/master/titanic_data.csv","titanic_data.csv")
TitanicData <- read_csv("titanic_data.csv")
```

CENSUS
```{r}
download.file("https://raw.githubusercontent.com/rashida048/Datasets/master/census.csv","census.csv")
CensusData <- read_csv("census.csv")
```

MOVIE
```{r}
download.file("https://raw.githubusercontent.com/rashida048/Datasets/master/movie_dataset.csv","movie_dataset.csv")
MovieData <- read_csv("movie_dataset.csv")
```

HOUSING PRICES
```{r}
download.file("https://raw.githubusercontent.com/rashida048/Datasets/master/home_data.csv","home_data.csv")
HomeData <- read_csv("home_data.csv")
```

CREDIT CARD FRUAD DATA
```{r}
download.file("https://raw.githubusercontent.com/rashida048/Datasets/master/fraud_data.csv","fraud_data.csv")
CCFruadData <- read_csv("fraud_data.csv")
```

